Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-09 base change.tex
Moved some material on (finite field) base change from base change.tex to shanghai.tex. Also, finished the proof of the proposition on connected commutative algebraic groups.

\section{Base change over the residue field}

When using character sheaves to study characters, it is useful to understand
how character sheaves behave under change of fields.
Suppose that $k'$ is a finite extension of $k$ of degree $n$, $G$ is an abelian group scheme
over $k$ as in Theorem \ref{thm:snake} and $G_{k'}$ is the base change of $G$ to $k'$.
If $\cs{L} = (\gcs{L}, \mu, F)$ is a character sheaf on $G$ we may define
a character sheaf $\nu(\cs{L}) = (\gcs{L}, \mu, F_{k'})$ on $G_{k'}$ by setting
\[
F_{k'} = F \circ \Frob_G^*(F) \circ \cdots \circ (\Frob_G^{n-1})^*(F).
\]
The commutativity of the diagram (CS.2) for $F_{k'}$ follows from the fact that $\Frob_{G_{k'}} = \Frob_G^n$.
Note that we may also think about the construction of $F_{k'}$ as taking the action of $\Weil{k}$ defined in Remark
\ref{rem:CS-action} and restricting it to $\Weil{k'}$.

\begin{proposition}
The following diagram commutes, where the bottom map is induced by the norm $G(k') \rightarrow G(k)$:
\[
\begin{tikzcd}
\CSiso{G} \rar{\nu} \dar{t_G} & \CSiso{G_{k'}} \dar{t_{G_{k'}}} \\
\Hom(G(k), \EEx) \rar{\Nm} & \Hom(G(k'), \EEx).
\end{tikzcd}
\]
\end{proposition}

>>>>

\section{Restriction of scalars}

### 2013-08-09 shanghai.tex
Moved some material on (finite field) base change from base change.tex to shanghai.tex. Also, finished the proof of the proposition on connected commutative algebraic groups.

\newcommand{\cdef}[1]{{{\color{blue}#1}\index{#1}}} 
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\obj}{{\text{obj}\, }}
\newcommand{\Gm}[1]{\mathbb{G}_{\hskip-1pt\textbf{m},#1}}
\newcommand{\GN}[1]{\mathswab{#1}}
\newcommand{\bGN}[1]{{\bar{\mathswab{#1}}}}
\newcommand{\TT}{\underline{T}}
\newcommand{\TL}{\underline{T_L}}
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
\newcommand{\CS}{{\mathcal{C\hskip-1.8pt S}}}
\newcommand{\GC}{{\mathcal{G\hskip-0.8pt C}}}
\newcommand{\CSiso}[1]{\CS(#1)_{/\textit{iso}}}
\newcommand{\CSbiso}[1]{\CS_0(#1)_{/\text{iso}}}
\newcommand{\CE}{{\mathcal{C\hskip-1.8pt E}}}
\newcommand{\Lgroup}[1]{{\,^L\hskip-1pt{#1}}}
\newcommand{\dualgroup}[1]{{\check{#1}}}
\newcommand{\Lang}{{\operatorname{Lang}}}
\newcommand{\image}{{\operatorname{im}}}
\newcommand{\Weil}[1]{\mathcal{W}_{#1}}
\newcommand{\Loc}{{\operatorname{Loc}}}

@@

@@

Let $\CS_0(G)$ be the category of pairs $(\cs{L}_0,\mu_0)$ 
where $\cs{L}_0$ is a constructible $\ell$-adic sheaf on $G$, 
which is a rank-$1$ local system on each connected component of $G$, 
equipped with an isomorphism $\mu_0 : m^* \cs{L}_0 \to \cs{L}_0 \boxtimes \cs{L}_0$ 
satisfying the analogue of Condition~\ref{CS.1} on $G$; 
morphisms in $\CS_f(G)$ are defined as in (the second part of) Condition~\ref{CS.2}. 
This is the category of \cdef{bounded character sheaves} on $G$. 
This is a rigid monoidal category in the obvious way, 
with duals given by $\cs{L}_0^\vee \ceq \sheafHom(\cs{L}_0,(\EE)_{G})$.

\begin{proposition}
$\CS_0(G)$ is (equivalent to) a full subcategory of $\CS(G)$.
\end{proposition}

\begin{proof}
 Let $b_G : {\bar G} \to G$ be the pull-back of $\Spec{\bFq} \to \Spec{\Fq}$ along $G\to \Spec{\Fq}$.
 Let $(\cs{L}_0,\mu_0)$ be a bounded character sheaf on $G$. 
 Then $\cs{L}_0$ is, in particular, an $\ell$-adic constructible sheaf on $G$. 
 Then $b_G^* \cs{L}_0$ comes equipped with an isomorphism 
 $F : \Frob_G^* b_G^*\cs{L}_0 \to b_G^* \cs{L}_0$; 
 moreover, the functor $\cs{L}_0 \mapsto (b_G^* \cs{L}_0,F)$  
 from $\ell$-adic constructible sheaves on $G$ to $\ell$-adic constructible sheaves on $G$
 is full and faithful; see \cite{SGA7.2}*{Expos\'e XIII} or even \cite{BBD}*{Prop. 5.2.1}. 
 This functor preserves local constancy, so takes local systems to local systems. 
 Set $\mu = b_{G\times G}^*\mu_0$; clearly, this satisfies Condition~\ref{CS.2} 
 with $b_G^*\cs{L}_0$ playing the role of $\gcs{L}$.
 Moreover, $F$ is compatible with $\mu$ in the sense of Condition~\ref{CS.3}.
 It follows that functor $(\cs{L}_0,\mu_0) \mapsto (b_G^*\cs{L}_0,b_{G\times G}^* \mu_0, F)$, denoted by  $B_G^*: \CS_0(G) \hookrightarrow \CS(G)$ below, is full and faithful.
\end{proof}

\Clifton{Address the elephant in the room: Is $B_G^*: \CS_0(G) \hookrightarrow \CS(G)$ an equivalence? I'm pretty sure I can show that the answer is `no', using trace of Frobenius, as you have suggested.}

@@

 %In this section we describe a full subcategory of $\CS_0(G)$
 %which will illuminate some features of $\CS_0(G)$ and also play a role in some arguments below.

To define the category of finite character sheaves we must work a bit harder.
Consider the following category $C(G)$:
an object in $C(G)$ is a pair $(f,\psi)$ where: 
$f : H\to G$ is an isogeny (a finite, surjective, etale, morphism of commutative group $\Fq$-schemes) 
for which the action of $\Gal(\bFq/\Fq)$ on the etale group scheme $\ker f$ is trivial, 
and $\psi : \ker f\to \Aut(V)$ is a representation on a finite-dimensional $\EE$-vector space $V$;
a map $(f,\psi) \to (f',\psi')$ in $C(G)$ is a pair $(\varphi,T)$ 
where $\varphi$ is a morphism of schemes such that $f' = f\circ \varphi$
and $T : V\to V'$ is a linear transformation equivariant 
for the action of $\ker f'$ on $V'$ by $\psi'$ 
and the action of $\ker f'$ on $V$ by $\psi \circ \varphi$.
%
We say that $(\varphi,T)$ in $C(G)$ is a weak isomorphism if $T$ is an isomorphism of vector spaces. 
Let $C(G)[W^{-1}]$ be the localization of $C(G)$ at weak isomorphisms.
Now, let $C_1(G)$ be the subcategory of $C(G)$ such that $\psi$ is a one-dimensional representation
and let $\CS_f(G)$ be the localisation of $C_1(G)$ at weak isomorphisms. 
This is the category of \cdef{finite character sheaves} on $G$.

@@

\begin{proposition}\label{prop:bounded}
 $\CS_f(G)$ is (equivalent to) a full subcategory of $\CS_0(G)$.
\end{proposition}

\begin{proof}
 To prove the proposition we exhibit a full and faithful monoidal functor $L_G : \CS_f(G) \to \CS_0(G)$.
 %
 We will do this by constructing a faithful monoidal functor $L : C_1(G) \to \CS_0(G)$ 
 such that ${\bar L}(\varphi,T)$ is an isomorphism precisely when $(\varphi,T)$ is a weak isomorphism. 
 Factoring along $C_1(G) \to \CS_0(G)$ will produce the full and faithful monoidal functor we seek.

@@

 One more ingredient is needed before we can exhibit the full functor $L_G: \CS_f(G) \to \CS_0(G)$ promised above: 
 we must put rigid monoidal category structures on $\CS_f(G)$ and $\CS_0(G)$.

@@

 If $(f,\psi)$ in an object in $C_1(G)$ then, as we have already seen, 
 $m^*L(f,\psi) \iso L(f,\psi) \boxtimes L(f,\psi)$ in $C_1(G\times G)$, canonically;
 let $\mu(f,\psi)$ be that isomorphism of $\ell$-adic sheaves on $G$,
 It follows immediately that the functor $L_G  \ceq  L_{W}\vert_{\CS_f(G)} : \CS_b(G) \to \CS_0(G)$, 
 given on objects by $(f,\psi) \mapsto (L(f,\psi), \mu(f,\psi))$,
 is a full and faithfull monoidal functor, completing the proof.
\end{proof}

@@

\begin{proposition}\label{prop:connected-1}
  If $G$ is a connected commutative algebraic group over $\Fq$ then 
  the full subcategories $\CS_f(G) \hookrightarrow \CS_0(G) \hookrightarrow \CS(G)$ 
  appearing in Section~\ref{sec:bounded} are equivalences 
  and (with reference to Section~\ref{sec:Frob})
  \[
  t_{G} : \CSiso{G} \to \Hom_\text{grp}(G(\Fq),\EEx)
  \]
  is an isomorphism of groups.
\end{proposition}

@@

  We may now lift the $\ell$-adic character $\rho_g : \Gal(\bFq/\Fq) \to \EEx$
  to an $\ell$-adic character $\pi_1(G,\bg) \to \EEx$ using the canonical topological group homomorphism
  $\pi_1(G,\bg) \to \Gal(\bFq/\Fq)$ determined by the structure map for $G$. But $\bg$ also
  determines an equivalence between the category of $\ell$-adic
  representations of $\pi_1(G,\bg)$ and $\ell$-adic local systems on $G$. Let
  $\cs{L}_0$ be a local system on $G$ in the isomorphism class
  determined by this $\ell$-adic character of $\pi_1(G,\bg)$.
  Then $b_G^*\cs{L}_0 \iso \gcs{L}$.
  Since $b_G^*$ is full and faithful (again, see \cite{SGA7.2}*{Expos\'e XIII} or even \cite{BBD}*{Prop. 5.2.1}),
  $
  b_{G\times G}^* : \Hom(m^*\cs{L}_0,\cs{L}_0\boxtimes\cs{L}_0) \to \Hom({\bar m}^*\gcs{L},\gcs{L}\boxtimes\gcs{L})
  $ 
  is a bijection 
  (hom taken in the categories on constructible $\ell$-adic sheaves on 
  $G\times G$ and ${\bar G}\times {\bar G}$ respectively, 
  in which $\ell$-adic local systems sit as full subcategories). 
  Let $\mu_0 : m^*\cs{L}_0 \to \cs{L}_0\boxtimes\cs{L}_0$ be the isomorphism matching 
  $\mu : {\bar m}^*\gcs{L} \to \gcs{L}\boxtimes\gcs{L}$, 
  the latter appearing in the definition of $\cs{L}$. 
  Then, as in Section~\ref{sec:bounded}, $(\cs{L}_0,\mu_0)$ is an object in $\CS_0(G)$ 
  and $\cs{L} = (\gcs{L},\mu,F)$ is isomorphic to $(b_G^*\cs{L}_0,b_{G\times G}^*\mu_0)$ in $\CS(G)$.
  Thus, the full and faithful functor $B_G^* : \CS_0(G) \to \CS(G)$ from Section~\ref{sec:bounded}
  is also essentially surjective, hence an equivalence.
  %(when $G$ is a connected, commutative algebraic group over $\Fq$).

  In Section~\ref{sec:finite} we saw that $\CS_f(G)$ is a full subcategory of $\CS_0(G)$;
  more precisely, we exhibited a full and faithful functor $L_G : \CS_f(G) \to \CS_0(G)$.
  We now show that this functor is essentially surjective 
  when $G$ is a connected, commutative algebraic group over $\Fq$.

@@

  We now show that $L_G(f,\psi) \iso (\cs{L}_0,\mu_0)$. 
  First, recall that $L_G(f,\psi) = (L(f,\psi), \mu(f,\psi))$. 
  By  \cite{Laumon}*{1.1.3.3}, $t_{L(f,\psi)}^{\Frob^n} = t_{\cs{L}}^{\Frob}\circ N_{\FF_{q^n}/\FF_q}$, for each positive integer $n$
  On the other hand, $t_{\cs{L}}^{\Frob}\circ N_{\FF_{q^n}/\FF_q} = 
  Using notation from \cite{Laumon}*{1.1},  this means $t_{\bullet L(f,\psi)} = t_{\bullet \cs{L}_0}$.
  It now \cite{Laumon}*{Th\'eor\`eme~1.1.2}, it follows that  $L(f,\psi) \iso \cs{L}_0$ in $\Loc(G)$ (admittedly, this is a big hammer).

   we simply apply the functions--sheaves dictionary
  (see \cite{Laumon}*{1.1.3} or \cite{SGA4.5}*{Sommes trigonométriques}) as follows.

  As we saw in Section~\ref{sec:finite}, the $\ell$-adic local system $L(f,\psi)$ on $G$ comes equipped with an isomorphism
  $m^* L(f,\psi) \iso L(f,\psi) \boxtimes L(f,\psi)$ in $\CS_b(G)$; 
  in the case at hand, this isomorphism is \cite[1.1.3.2]{Laumon}.
  By \cite[1.1.3.2]{Laumon},

  Since we have already seen (in Section~\ref{sec:finite}) that $L_G$ is full and faithful, 
  it will then follow that $L_G$ is an equivalence.

  \todo{Clifton is here. I will use \cite{Kamgarpour} to finish the proof.}

  We have now shown that $b_G^* : \CS_0(G) \to \CS(G)$ is an equivalence
  of rigid monoidal categories. It follows immediately that $b_G^*$
  determines an isomorphism
  \[
  b_G^* : \CSbiso{G} \mathop{\longrightarrow}^{\iso} \CSiso{G}
  \]
  of the Picard groups determined by these rigid monoidal categories.
  \Clifton{\textdbend This last sentence glosses over some subtle points;
    I might need to address them in the `two perspectives' section.}

  To finish the proof observe that the functions--sheaves dictionary,
  as it pertains to characters, shows that
  \[
  t_{G} : \CSbiso{G} \to \Hom_\text{grp}(G(\Fq),\EEx)
  \]
  is an isomorphism of groups; see \cite{Laumon}*{1.1.3} or \cite{SGA4.5}*{Sommes trigonométriques}.
\end{proof}

\section{Commutative etale group schemes}

@@

%\input{etale-2}
%Define $\rho \colon \Hh^1(\Weil{\Fq}, \Hom(G(\bFq), \EEx))
% \rightarrow \Hom(G(\Fq), \EEx)$ by $\rho(\phi)(g) =
% \phi(\Frob)(g)$.

@@

\begin{proof}
  Because $G$ is smooth over $\Fq$, the identity component $G^0$ is a
  geometrically connected algebraic group over $\Fq$, clopen in $G$
  \cite{vdG&M}*{3.17}. By Proposition~\ref{prop:connected-1}, every
  character sheaf $\cs{E} = (\gcs{E},\mu^0,F^0)$ on $G^0$ descends to
  a line bundle on $G^0$.

@@

\begin{proof}
  Consider the short exact sequence in the category of smooth group
  schemes, locally of finite type over $\Fq$, defining the component
  group scheme for $G$:
  \begin{equation}\label{eq:pi0}
  \begin{tikzcd}
    1 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
  \end{tikzcd}
  \end{equation}
  Let
  \[
  \begin{tikzcd}[row sep=30]
    {}& \ker t_{\pi_0(G)} \dar & \arrow[dashed]{d} \ker t_{G} & \ker t_{G^0} \dar & \\
    0 \rar & \CSiso{\pi_0(G)} \rar \dar{t_{\pi_0(G)}}
    & \CSiso{G} \rar \dar{t_{G}} & \CSiso{G^0} \rar \dar{t_{G^0}} & 0\\
    0 \rar & \Hom(\pi_0(G)(\Fq),\EEx) \rar \dar
    & \arrow[dashed]{d} \Hom(G(\Fq),\EEx) \rar & \Hom(G^0(\Fq),\EEx) \rar \dar & 0\\
    & \coker t_{\pi_0(G)} & \coker t_{G} &  \coker t_{G^0} &
  \end{tikzcd}
  \]
  be the commutative diagram of abelian groups obtained by applying
  Proposition~\ref{prop:pullback} to \eqref{eq:pi0}.  By
  Proposition~\ref{prop:connected-1}, $\ker t_{G^0} =0$ and $\coker t_{G^0}=0$.
  By Propositions~\ref{prop:etale} and \ref{prop:X}, $\ker t_{\pi_0(G)}=0$
  and $\coker t_{\pi_0(G)}=0$.

>>>>

\newcommand{\cdef}[1]{{{\color{blue}#1}\index{#1}}} 
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\obj}{{\text{obj}\, }}
\newcommand{\Gm}[1]{\mathbb{G}_{\hskip-1pt\textbf{m},#1}}
\newcommand{\GN}[1]{\mathswab{#1}}
\newcommand{\bGN}[1]{{\bar{\mathswab{#1}}}}
\newcommand{\TT}{\underline{T}}
\newcommand{\TL}{\underline{T_L}}
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
\newcommand{\CS}{{\mathcal{C\hskip-1.8pt S}}}
\newcommand{\CSb}{{\mathcal{C\hskip-1.8pt S}_0}}
\newcommand{\CSf}{{\mathcal{C\hskip-1.8pt S}_{f}}}
\newcommand{\CSiso}[1]{\CS(#1)_{/\textit{iso}}}
\newcommand{\CSbiso}[1]{\CSb(#1)_{/\text{iso}}}
\newcommand{\CSfiso}[1]{\CSf(#1)_{/\text{iso}}}
\newcommand{\CE}{{\mathcal{C\hskip-1.8pt E}}}
\newcommand{\Lgroup}[1]{{\,^L\hskip-1pt{#1}}}
\newcommand{\dualgroup}[1]{{\check{#1}}}
\newcommand{\Lang}{{\operatorname{Lang}}}
\newcommand{\image}{{\operatorname{im}}}
\newcommand{\Weil}[1]{\mathcal{W}_{#1}}
\newcommand{\Loc}{{\operatorname{Loc}}}

@@

\section{Base change over the residue field}\label{sec:basechange}

When using character sheaves to study characters, it is useful to understand
how character sheaves behave under change of fields.
Suppose that $k'$ is a finite extension of $k$ of degree $n$, $G$ is an abelian group scheme
over $k$ as in Theorem \ref{thm:snake} and $G_{k'}$ is the base change of $G$ to $k'$.
If $\cs{L} = (\gcs{L}, \mu, F)$ is a character sheaf on $G$ we may define
a character sheaf $\nu(\cs{L}) = (\gcs{L}, \mu, F_{k'})$ on $G_{k'}$ by setting
\[
F_{k'} = F \circ \Frob_G^*(F) \circ \cdots \circ (\Frob_G^{n-1})^*(F).
\]
The commutativity of the diagram (CS.2) for $F_{k'}$ follows from the fact that $\Frob_{G_{k'}} = \Frob_G^n$.
Note that we may also think about the construction of $F_{k'}$ as taking the action of $\Weil{k}$ defined in Remark
\ref{rem:CS-action} and restricting it to $\Weil{k'}$.

\begin{proposition}\label{prop:basechange}
The following diagram commutes, where the bottom map is induced by the norm $G(k') \rightarrow G(k)$:
\[
\begin{tikzcd}
\CSiso{G} \rar{\nu} \dar{t_G} & \CSiso{G_{k'}} \dar{t_{G_{k'}}} \\
\Hom(G(k), \EEx) \rar{\Nm} & \Hom(G(k'), \EEx).
\end{tikzcd}
\]
\end{proposition}

\begin{proof}
\todo{This proof is missing.}
\end{proof}

@@

Let $\CSb(G)$ be the category of pairs $(\cs{L}_0,\mu_0)$ 
where $\cs{L}_0$ is a constructible $\ell$-adic sheaf on $G$, 
which is a rank-$1$ local system on each connected component of $G$, 
equipped with an isomorphism $\mu_0 : m^* \cs{L}_0 \to \cs{L}_0 \boxtimes \cs{L}_0$ 
satisfying the analogue of Condition~\ref{CS.1} on $G$; 
morphisms in $\CSf(G)$ are defined as in (the second part of) Condition~\ref{CS.2}. 
This is the category of \cdef{bounded character sheaves} on $G$. 
This is a rigid monoidal category in the obvious way, 
with duals given by $\cs{L}_0^\vee \ceq \sheafHom(\cs{L}_0,(\EE)_{G})$.

\begin{proposition}
$\CSb(G)$ is (equivalent to) a full subcategory of $\CS(G)$.
\end{proposition}

\begin{proof}
 Let $b_G : {\bar G} \to G$ be the pull-back of $\Spec{\bFq} \to \Spec{\Fq}$ along $G\to \Spec{\Fq}$.
 Let $(\cs{L}_0,\mu_0)$ be a bounded character sheaf on $G$. 
 Then $\cs{L}_0$ is, in particular, an $\ell$-adic constructible sheaf on $G$. 
 Then $b_G^* \cs{L}_0$ comes equipped with an isomorphism 
 $F : \Frob_G^* b_G^*\cs{L}_0 \to b_G^* \cs{L}_0$; 
 moreover, the functor $\cs{L}_0 \mapsto (b_G^* \cs{L}_0,F)$  
 from $\ell$-adic constructible sheaves on $G$ to $\ell$-adic constructible sheaves on $G$
 is full and faithful; see \cite{SGA7.2}*{Expos\'e XIII} or even \cite{BBD}*{Prop. 5.2.1}. 
 This functor preserves local constancy, so takes local systems to local systems. 
 Set $\mu = b_{G\times G}^*\mu_0$; clearly, this satisfies Condition~\ref{CS.2} 
 with $b_G^*\cs{L}_0$ playing the role of $\gcs{L}$.
 Moreover, $F$ is compatible with $\mu$ in the sense of Condition~\ref{CS.3}.
 It follows that functor $(\cs{L}_0,\mu_0) \mapsto (b_G^*\cs{L}_0,b_{G\times G}^* \mu_0, F)$, denoted by  $B_G^*: \CSb(G) \hookrightarrow \CS(G)$ below, is full and faithful.
\end{proof}

@@

 %In this section we describe a full subcategory of $\CSb(G)$
 %which will illuminate some features of $\CSb(G)$ and also play a role in some arguments below.

To define the category of finite character sheaves we must work a bit harder.
Consider the following category $C(G)$:
an object in $C(G)$ is a pair $(f,\psi)$ where: 
$f : H\to G$ is an isogeny (a finite, surjective, etale, morphism of commutative group $\Fq$-schemes) 
for which the action of $\Gal(\bFq/\Fq)$ on the etale group scheme $\ker f$ is trivial, 
and $\psi : \ker f\to \Aut(V)$ is a representation on a finite-dimensional $\EE$-vector space $V$;
a map $(f,\psi) \to (f',\psi')$ in $C(G)$ is a pair $(\varphi,T)$ 
where $\varphi$ is a morphism of schemes such that $f' = f\circ \varphi$
and $T : V\to V'$ is a linear transformation equivariant 
for the action of $\ker f'$ on $V'$ by $\psi'$ 
and the action of $\ker f'$ on $V$ by $\psi \circ \varphi$.
%
We say that $(\varphi,T)$ in $C(G)$ is a weak isomorphism if $T$ is an isomorphism of vector spaces. 
Let $C(G)[W^{-1}]$ be the localization of $C(G)$ at weak isomorphisms.
Now, let $C_1(G)$ be the subcategory of $C(G)$ such that $\psi$ is a one-dimensional representation
and let $\CSf(G)$ be the localisation of $C_1(G)$ at weak isomorphisms. 
This is the category of \cdef{finite character sheaves} on $G$.

@@

\begin{proposition}\label{prop:bounded}
 $\CSf(G)$ is (equivalent to) a full subcategory of $\CSb(G)$.
\end{proposition}

\begin{proof}
 To prove the proposition we exhibit a full and faithful monoidal functor $L_G : \CSf(G) \to \CSb(G)$.
 %
 We will do this by constructing a faithful monoidal functor $L : C_1(G) \to \CSb(G)$ 
 such that ${\bar L}(\varphi,T)$ is an isomorphism precisely when $(\varphi,T)$ is a weak isomorphism. 
 Factoring along $C_1(G) \to \CSb(G)$ will produce the full and faithful monoidal functor we seek.

@@

 One more ingredient is needed before we can exhibit the full functor $L_G: \CSf(G) \to \CSb(G)$ promised above: 
 we must put rigid monoidal category structures on $\CSf(G)$ and $\CSb(G)$.

@@

 If $(f,\psi)$ in an object in $C_1(G)$ then, as we have already seen, 
 $m^*L(f,\psi) \iso L(f,\psi) \boxtimes L(f,\psi)$ in $C_1(G\times G)$, canonically;
 let $\mu(f,\psi)$ be that isomorphism of $\ell$-adic sheaves on $G$,
 It follows immediately that the functor $L_G  \ceq  L_{W}\vert_{\CSf(G)} : \CS_b(G) \to \CSb(G)$, 
 given on objects by $(f,\psi) \mapsto (L(f,\psi), \mu(f,\psi))$,
 is a full and faithfull monoidal functor, completing the proof.
\end{proof}

@@

\begin{proposition}\label{prop:connected}
  If $G$ is a connected commutative algebraic group over $\Fq$ then 
  the full subcategories $\CSf(G) \hookrightarrow \CSb(G) \hookrightarrow \CS(G)$ 
  appearing in Section~\ref{sec:bounded} are equivalences 
  and (with reference to Section~\ref{sec:Frob})
  \[
  t_{G} : \CSiso{G} \to \Hom_\text{grp}(G(\Fq),\EEx)
  \]
  is an isomorphism of groups.
\end{proposition}

@@

  We may now lift the $\ell$-adic character $\rho_g : \Gal(\bFq/\Fq) \to \EEx$
  to an $\ell$-adic character $\pi_1(G,\bg) \to \EEx$ using the canonical topological group homomorphism
  $\pi_1(G,\bg) \to \Gal(\bFq/\Fq)$ determined by the structure map for $G$. But $\bg$ also
  determines an equivalence between the category of $\ell$-adic
  representations of $\pi_1(G,\bg)$ and $\ell$-adic local systems on $G$. Let
  $\cs{L}_0$ be a local system on $G$ in the isomorphism class
  determined by this $\ell$-adic character of $\pi_1(G,\bg)$.
  Then $b_G^*\cs{L}_0 \iso \gcs{L}$.
  Since $b_G^*$ is full and faithful (again, see \cite{SGA7.2}*{Expos\'e XIII} or even \cite{BBD}*{Prop. 5.2.1}),
  $
  b_{G\times G}^* : \Hom(m^*\cs{L}_0,\cs{L}_0\boxtimes\cs{L}_0) \to \Hom({\bar m}^*\gcs{L},\gcs{L}\boxtimes\gcs{L})
  $ 
  is a bijection 
  (hom taken in the categories on constructible $\ell$-adic sheaves on 
  $G\times G$ and ${\bar G}\times {\bar G}$ respectively, 
  in which $\ell$-adic local systems sit as full subcategories). 
  Let $\mu_0 : m^*\cs{L}_0 \to \cs{L}_0\boxtimes\cs{L}_0$ be the isomorphism matching 
  $\mu : {\bar m}^*\gcs{L} \to \gcs{L}\boxtimes\gcs{L}$, 
  the latter appearing in the definition of $\cs{L}$. 
  Then, as in Section~\ref{sec:bounded}, $(\cs{L}_0,\mu_0)$ is an object in $\CSb(G)$ 
  and $\cs{L} = (\gcs{L},\mu,F)$ is isomorphic to $(b_G^*\cs{L}_0,b_{G\times G}^*\mu_0)$ in $\CS(G)$.
  Thus, the full and faithful functor $B_G^* : \CSb(G) \to \CS(G)$ from Section~\ref{sec:bounded}
  is also essentially surjective, hence an equivalence.
  %(when $G$ is a connected, commutative algebraic group over $\Fq$).

  In Section~\ref{sec:finite} we saw that $\CSf(G)$ is a full subcategory of $\CSb(G)$;
  more precisely, we exhibited a full and faithful functor $L_G : \CSf(G) \to \CSb(G)$.
  We now show that this functor is essentially surjective 
  when $G$ is a connected, commutative algebraic group over $\Fq$.

@@

  To show that $L_G(f,\psi) \iso (\cs{L}_0,\mu_0)$ we simply apply the functions--sheaves dictionary
  as follows ({\it cf} \cite{Laumon}*{1.1.3} or, ultimately, \cite{SGA4.5}*{Sommes trigonométriques}).
  First, recall that $L_G(f,\psi) = (L(f,\psi), \mu(f,\psi))$. 
  By  \cite{Laumon}*{1.1.3.3}, $t_{L(f,\psi)}^{\Frob^n} = t_{\cs{L}}^{\Frob}\circ N_{\FF_{q^n}/\FF_q}$, for each positive integer $n$.
  On the other hand, $t_{\cs{L}}^{\Frob} \circ N_{\FF_{q^n}/\FF_q} = t_{\cs{L}}^{\Frob^n}$, for each positive integer $n$, 
  by Proposition~\ref{prop:basechange}.
  %Using notation from \cite{Laumon}*{1.1}, this means $t_{\bullet L(f,\psi)} = t_{\bullet \cs{L}_0}$.
  Using \cite{Laumon}*{Th\'eor\`eme~1.1.2}, it follows that  $L(f,\psi) \iso \cs{L}_0$ in $\Loc(G)$.
  It is now clear that $(L(f,\psi),\mu(f,\psi)) \iso (\cs{L}_0,\mu_0)$ in $\CSb(G)$.
  This completes the proof that $L_G : \CSf(G) \to \CS(G)$ is essentially surjective.
  Since we have already seen (in Section~\ref{sec:finite}) that $L_G$ is full and faithful, 
  it follows that $L_G$ is an equivalence.

  To finish the proof of Proposition~\ref{prop:connected}, it suffices to observe that the group homomorphism
  \[
  t_{G} : \CSiso{G} \to \Hom_\text{grp}(G(\Fq),\EEx)
  \]
  is surjective because, as we have seen, 
  each character $\chi \in \Hom_\text{grp}(G(\Fq),\EEx)$ determines a finite character sheaf $(\Lang,\chi)$
  such that the trace of Frobenius of the character sheaf $B_G^* (L_G(\Lang,\chi))$ is $\chi$.
\end{proof}

\section{Character sheaves on commutative etale group schemes}

@@

\section{On the cohomology of the dual of an etale group scheme}
%\input{etale-2}
%Define $\rho \colon \Hh^1(\Weil{\Fq}, \Hom(G(\bFq), \EEx))
% \rightarrow \Hom(G(\Fq), \EEx)$ by $\rho(\phi)(g) =
% \phi(\Frob)(g)$.

Although this section does not mention character sheaves, 
we will need Proposition~\ref{prop:X} to prove our main result on character sheaves, 
Theorem~\ref{thm:snake}.

@@

\begin{proof}
  Because $G$ is smooth over $\Fq$, the identity component $G^0$ is a
  geometrically connected algebraic group over $\Fq$, clopen in $G$
  \cite{vdG&M}*{3.17}. By Proposition~\ref{prop:connected}, every
  character sheaf $\cs{E} = (\gcs{E},\mu^0,F^0)$ on $G^0$ descends to
  a line bundle on $G^0$.

@@

\begin{proof}
  Consider the short exact sequence in the category of smooth group
  schemes, locally of finite type over $\Fq$, defining the component
  group scheme for $G$:
  \begin{equation}\label{eq:pi0}
  \begin{tikzcd}
    1 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
  \end{tikzcd}
  \end{equation}
  Let
  \[
  \begin{tikzcd}[row sep=30]
    {}& \ker t_{\pi_0(G)} \dar & \arrow[dashed]{d} \ker t_{G} & \ker t_{G^0} \dar & \\
    0 \rar & \CSiso{\pi_0(G)} \rar \dar{t_{\pi_0(G)}}
    & \CSiso{G} \rar \dar{t_{G}} & \CSiso{G^0} \rar \dar{t_{G^0}} & 0\\
    0 \rar & \Hom(\pi_0(G)(\Fq),\EEx) \rar \dar
    & \arrow[dashed]{d} \Hom(G(\Fq),\EEx) \rar & \Hom(G^0(\Fq),\EEx) \rar \dar & 0\\
    & \coker t_{\pi_0(G)} & \coker t_{G} &  \coker t_{G^0} &
  \end{tikzcd}
  \]
  be the commutative diagram of abelian groups obtained by applying
  Proposition~\ref{prop:pullback} to \eqref{eq:pi0}.  By
  Proposition~\ref{prop:connected}, $\ker t_{G^0} =0$ and $\coker t_{G^0}=0$.
  By Propositions~\ref{prop:etale} and \ref{prop:X}, $\ker t_{\pi_0(G)}=0$
  and $\coker t_{\pi_0(G)}=0$.
